### 2019-06-03 article/article.tex
Working on article, huge enhancements of some metrics

\documentclass{jdmdh}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{pgfplots}

\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French}
\author[1]{Thibault Clérice}
\affil[1]{École nationale des Chartes, France} 
\affil[2]{Université Lyon 3, France}

\corrauthor{Thibault Clérice}{thibault.clerice@chartes.psl.eu}

\pgfplotsset{compat=1.15}
\begin{document}

\maketitle

\abstract{Tokenization of modern and old Western European languages seems to be fairly simple as it stands on the presence mostly of markers such as spaces and punctuation. Although, when dealing with old sources like manuscript written in \textit{scripta continua}, (1) such markers are mostly absent, (2) spelling variation and rich morphology makes dictionary based approaches difficult. We show that applying convolutional encoding to characters followed by linear categorization to word-boundary or in-word-sequence can be used to tokenize such inputs. Additionally, we release a software with a rather simple interface for tokenizing one's corpus.}

\keywords{convolutional network; scripta continua; tokenization; Old French; word segmentation}

\section{Introduction}

% To Read : Stutzmann article.
% Parler plus de la variation orthographique de langue

Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando". \textit{Scripta continua} is a writing phenomenon where words would not be separated by spaces and it appears to have disappeared around the 8th century (see \citet{zanna1998lecture}). Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century. In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{4-lines-p0215.png}

  \caption{ 4 lines from fol.103rb Manuscript fr. 412, Bibliothèque nationale de France.  Red lines indicate word boundaries}
  \label{fig:4lines}
\end{figure}

We must stress in this study that the difficulty that we face is different for \textit{scripta continua} than the ones researchers face languages such as Chinese for which an already impressive amount of work has been done as it. Indeed, Chinese word segmentation has lately been driven by deep learning methods, specifically ones based on \textit{sequence to sequence translations}: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses BiDirectional GRU and CRF. Actually, meanwhile redacting this article and producing the code-base, \citet{huang2019realistic} took the same approach of encoding to linear classification to both word boundary (WB) and word content (WC) for Chinese word segmentation.

Indeed, while Chinese's issue seems to lie in the decomposition of relatively fix characters, Old French or medieval latin present heavy variation of spelling. In \citet{camps_pandora}, Camps notes, in the same corpus, the existence of not less than 29 spelling of the word \textit{cheval} (horse in Old and Modern French) whose apparition counts span from 3907 to 1\footnote{These are \textit{cheval}, \textit{chevaus}, \textit{cheual}, \textit{ceval}, \textit{chevals}, \textit{cevaus}, \textit{chival}, \textit{ceual}, \textit{cheuaus}, \textit{cevals}, \textit{chaval}, \textit{chivaus}, \textit{chiual}, \textit{chevas}, \textit{cheuals}, \textit{chiuaus}, \textit{ceuaus}, \textit{chevaul}, \textit{chiuau}, \textit{chivals}, \textit{chevau}, \textit{kevaus}, \textit{chavaus}, \textit{cheuas}, \textit{keval}, \textit{cheua}, \textit{cheuau}, \textit{cheva}, \textit{chiuals}}. This  makes a dictionary approach rather difficult as it would rely on a high number of different spelling and makes the computation highly complex.

\section{Description and evaluation}

\subsection{Architecture}

\subsubsection{Encoding of input and decoding}

The model is based on traditional text input encoding where each character is transcoded to an index. Output of the model is a mask that needs to be applied to the input: in the mask, characters are classified either as word boundary or word content (\textit{cf.} Table \ref{lst:input_output_example}.

\begin{table}
\centering
\begin{tabular}{@{}ll@{}}
\hline
                       & \textbf{Sample}           \\  \hline
\textbf{Input  String} & \texttt{Ladamehaitees'enparti}     \\
\textbf{Mask   String} & \texttt{xSxxxSxxxxxSxxxSxxxxS}     \\
\textbf{Output String} & \texttt{La dame haitee s'en parti} \\ \hline
\end{tabular}
  \caption{Input, mask and human-readable output generated by the model. x are WC and S are WB}
  \label{lst:input_output_example}
\end{table}

For evaluation purposes, and to reduce the number of input classes, we propose two options for data transcoding: a lower-case normalization and a "reduction to the ASCII character set" feature (fr. \ref{fig:normalization}). On this point, a lot of issues were found with transliteration of medieval paelographic characters that were part of the original datasets, as they are badly interpreted by the \texttt{unidecode} python package. Indeed, \texttt{unidecode} will simply remove characters it does not understand. I built a secondary package named \texttt{mufidecode} (\citet{thibault_clerice_2019_3237731}) which precedes unidecode equivalency tables when the data is known of the Medieval Unicode Font Initiative (MUFI, \citet{mufi}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{carbon.png}
  \caption{Different possibilities of pre-processing. The option with join=False was kept, as it keeps abbreviation marked as single characters. Note how \texttt{unidecode} loses the P WITH BAR}
  \label{fig:normalization}
\end{figure}

\subsubsection{Model}

Aside from normalizations of the input and output, three different structure of models were tested. Every model is composed by one encoder described below and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary). For final scores, SOS, EOS and PAD were ignored.

The encoders are the following (configurations in parenthesis):

\begin{itemize}
  \item LSTM encoder with hidden cell (Embedding (512), Dropout(0.5), Hidden Dimension (512), Layers(10))
  \item Convolutional (CNN) encoder with position embeddings (Embedding (256), Embedding(Maximum Sentence Size=150), Kernel Size (5), Dropout(0.25), Layers (10))
  \item Convolutional (CNN) encoder without position embeddings (Embedding (256), Kernel Size (5), Dropout(0.25), Layers (10))
\end{itemize}

\subsection{Evaluation}

\subsubsection{Datasets}

Datasets are transcription from manuscripts with unresolved abbreviation coming from different projects. The \textbf{Old French} is based on \citet{8269990}, \citet{pinche:hal-01628533}, \citet{jean_baptiste_camps_2019_2630574}, \citet{bfmmss}, and \citet{tnah_transcription}. It contains

\begin{itemize}
    \item 193,734 training examples;
    \item 23,581 validation examples;
    \item 25,512 test examples
\end{itemize}

The input was generated by grouping at least 2 words and a maximum of 8 words together per sample. On a probability of 0.2, noise character could be added (noise character was set to DOT ('.')) and some words were kept randomly from a sample to another on a probability of 0.3 and a maximum number of word kept of 1. If a minimum size of 7 characters was not met in the input sample, another word would be added to the chain. A maximum input size of 100 was kept. The results corpora should be varied in sizes as shown by \ref{fig:word_sizes}. The corpora is composed by 193 different characters when not normalized, in which some MUFI characters appears few hundred times \ref{tab:mufi_examples}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{length.png}
  \caption{Distribution of word size over the train, dev and test corpora}
  \label{fig:word_sizes}
\end{figure}

\begin{table}
\begin{tabular}{llll}
\hline
                                                   & Train dataset & Dev dataset & Test dataset \\ \hline
TIRONIAN SIGN ET                                   & 4367          & 541         & 539          \\
CON                             & 508           & 70          & 76           \\
P WITH STROKE THROUGH DESCENDER & 580           & 69          & 84           \\ \hline
\end{tabular}
  \caption{Examples of some MUFI characters distributions}
  \label{tab:mufi_examples}
\end{table}

\subsubsection{Results}

The training parameters was 0.00005 in learning rate for each CNN model and 0.001 for the LSTM one, and 64 in batch sizes. Training reached a plateau fairly quickly for each model (\textit{cf.} \ref{fig:loss}). Each model except LSTM reached a really low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores})

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
          width=\linewidth, % Scale the plot to \linewidth
          grid=major,
          grid style={dashed,gray!30},
          xlabel=Epoch, % Set the labels
          ylabel=Accuracy,
          legend style={at={(0.5,-0.2)},anchor=north},
          x tick label style={rotate=90,anchor=east}
        ]
        \addplot table[x=Epoch,y=CNN 1,col sep=comma] {accuracies.csv};
        \addplot table[x=Epoch,y=CNN L,col sep=comma] {accuracies.csv};
        \addplot table[x=Epoch,y=CNN w/o P,col sep=comma] {accuracies.csv};
        \addplot table[x=Epoch,y=CNN N,col sep=comma] {accuracies.csv};
        \addplot table[x=Epoch,y=CNN L N,col sep=comma] {accuracies.csv};
        \legend{CNN, CNN L, CNN P, CNN N, CNN L N}
      \end{axis}
    \end{tikzpicture}
    \caption{Training Loss (Cross-entropy) until plateau was reached. N = normalized, L = Lower, P = no position embedding. LSTM was removed as it did not go below 0.65}
  \label{fig:loss}
  \end{center}
\end{figure}

\begin{table}
\centering
\begin{tabular}{lllll}
\hline
Model   & Accuracy & Precision & Recall & FScore \\ \hline
CNN     & 0.991    & 0.985     & 0.990  & 0.987  \\
CNN L   & 0.991    & 0.979     & 0.990  & 0.985  \\
CNN P   & \textbf{0.993}    & \textbf{0.990}& \textbf{0.991}  & \textbf{0.990}  \\
CNN N   & 0.991    & 0.987     & 0.988  & 0.988  \\
CNN L N & 0.992    & 0.988     & 0.989  & 0.988  \\
LSTM    & 0.741    & 0.184     & 0.500  & 0.269  \\ \hline
\end{tabular}
\caption{Scores over the test dataset. N = normalized, L = Lower, P = no position embedding.}
\label{tab:scores}
\end{table}

\subsubsection{Example of outputs}

The following inputs has been tagged with the CNN P model. Batch are constructed around the regex \texttt{\\W} with package \texttt{regex}. This explains why inputs such as \texttt{".i."} are automatically tagged as \texttt{" . i . "} by the tool. The input was stripped of its spaces before tagging, we only show the ground truth by commodity.

\begin{itemize}
    \item \texttt{Truth :} Aies joie et leesce en ton cuer car tu auras une fille qui aura .i. fil qui sera de molt grant merite devant Dieu et de grant los entre les homes.
    \item \texttt{Input :} Aiesjoieetleesceentoncuercartuaurasunefillequiaura.i.filquiserademoltgrantmeritedevantDieuetdegrantlosentreleshomes.
    \item \texttt{CNN :} Aiesjoie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
    \item \texttt{CNN lower:}	Aies joie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
    \item \texttt{CNN without position:} Aiesjoie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
    \item \texttt{CNN Normalize:}	Aies joie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
\end{itemize}

\subsection{Discussion}

We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view

% "The purpose of the discussion is to interpret and describe the significance of your findings in light of what was already known about the research problem being investigated and to explain any new understanding or insights that emerged as a result of your study of the problem. The discussion will always connect to the introduction by way of the research questions or hypotheses you posed and the literature you reviewed, but the discussion does not simply repeat or rearrange the first parts of your paper; the discussion clearly explain how your study advanced the reader's understanding of the research problem from where you left them at the end of your review of prior research."

\subsection{Conclusion}

While
% The conclusion is intended to help the reader understand why your research should matter to them after they have finished reading the paper. A conclusion is not merely a summary of the main topics covered or a re-statement of your research problem, but a synthesis of key points and, if applicable, where you recommend new areas for future research. For most college-level research papers, one or two well-developed paragraphs is sufficient for a conclusion, although in some cases, three or more paragraphs may be required.

\subsection{Acknowledgement}

Boudams has been made possible by two open-source repositories from which I learned and copied bits of implementation of certain modules and without which none of this paper would have been possible: \citet{enrique_manjavacas_2019_2654987} and \citet{bentrevett}. This tool was originally intended for post-processing OCR for the presentation \citet{pinchecampsclerice} at DH2019 in Utrecht.

\bibliographystyle{plainnat}
\bibliography{article}

\appendix\footnotesize

\section{Annex 1 : Confusion of CNN without position embeddings}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{confusion.png}
  \caption{Confusion matrix of the CNN model without position embedding}
  \label{fig:confusion_matrix}
\end{figure}

\section{Annex 2}
Cras tristique vel nisi at aliquet. Proin egestas erat sit amet velit lobortis imperdiet. Integer et arcu sapien. Etiam id blandit
sapien. Nam tempus lacus ac massa semper, vel laoreet turpis rutrum. Mauris eget nibh vitae justo porta imperdiet sed vel
ligula. In imperdiet, augue vel condimentum convallis, neque augue imperdiet neque, eget dapibus nunc mauris ultricies
tortor. Nam eget nunc egestas, blandit lectus non, aliquam nunc. Cras sed quam vitae arcu ornare lobortis. Ut ut lacus
hendrerit, convallis orci sit amet, commodo nunc. Pellentesque eget tincidunt tortor. Nunc ornare molestie mauris id vehicula.
Suspendisse pharetra tortor metus, sit amet fermentum tellus vehicula ut.

\end{document}